---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# malariaEquilibrium

[![Travis build status](https://travis-ci.org/mrc-ide/malariaEquilibrium.svg?branch=master)](https://travis-ci.org/mrc-ide/malariaEquilibrium)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mrc-ide/malariaEquilibrium?branch=master&svg=true)](https://ci.appveyor.com/project/mrc-ide/malariaEquilibrium)

Often we are interested in the state of a malaria transmission model at
equibrium. However, some models (e.g. the Griffin et al. 2014 model) are quite
complex, and can result in equilibrium solutions that are fairly in-depth. In
these situations it is useful to have a "canonical" equilibrium solution that is
tried and tested, and can be used reliably by multiple users. This package aims
to be a place for hosting these canonical solutions, and for storing useful
tests, checks and plotting functions for exploring model predictions at
equilibrium.

## How to interact with this repos

If you have a new model with an equilibrium solution that you would like to host
here, then please go ahead and do so *in a new branch*. The master branch, on
the other hand, should always correspond to the current most active model used
by the malaria group (currently the Griffin et al. 2014 model). This should
generally be left alone unless you know what you're doing.

Any changes that you propose to branches that you did not create yourself (e.g.
the master branch) should be done through pull requests, rather than editing
directly. This gives others the opportunity to review your changes before they
are implemented in code.

NB. These stipulations are in place because there are other R packages that test
against the equilibrium solutions from this package, and so any changes here may
cause other packages to fail if they are not know about beforehand.

## Installation

You can install from the master branch of malariaEquilibrium with:

``` r
# install.packages("devtools")
devtools::install_github("mrc-ide/malariaEquilibrium")
```

Or from a specific branch with:

``` r
# install.packages("devtools")
devtools::install_github("mrc-ide/malariaEquilibrium", ref = "mybranch")
```

Finally, don't forget to load the package with:

```{r}
library(malariaEquilibrium)
```

## Example

### Working with parameters

We start by loading a set of model parameters with the `load_parameter_set()`
function. This loads an object from the *inst/extdata* folder of the package.
Alternatively you can load your own parameter set from file.

```{r}
myparams <- load_parameter_set("Jamie_parameters.rds")
```

Parameters are of class `model_params`, which is essentially just an ordinary R
list with a few custom methods. For example, when we print we see the following:

```{r}
myparams
```

We can modify parameter values exactly as we would with a named list:

```{r}
myparams$eta <- 0.1
head(myparams)
```

### Getting the equilibrium solution

Now we can calculate the equilibrium solution using the `human_equilibrium()`
function. This assumes a fixed EIR and treatment rate (`ft`), and takes model
parameters as input. It calculates the equilibrium solution for a defined age
range:

```{r}
eq <- human_equilibrium(EIR = 10, ft = 0.2, p = myparams, age = 0:10)
```

The output elements are "states" and "FOIM". The former is a matrix with age
brackets in rows, and model states or other downstream calculations in columns.
For example, the column "S" gives the equilibrium proportion of humans in the
susceptible state, while the column "pos_M" gives the proportion of hosts
positive by microscopy in this bracket.

```{r}
head(eq$states)
```

The "FOIM" element gives the onward force of infection from humans to
mosquitoes, which is a single number:

```{r}
eq$FOIM
```
